---
title: "Homework 1"
author: "Malvika Venkataraman"
date: '2021-09-22'
output:
  html_document:
    df_print: paged
---

```{r, include = FALSE}
library(tidyverse)
```

## Problem 1

### set up data frame
```{r}
set.seed(1)
data_frame = tibble(
  vec_numeric = rnorm(10),
  vec_logical = vec_numeric > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(c("single", "married", "divorced", "married", "divorced", "single", "single", "divorced", "married", "single"))
)
```
### take means

```{r}
mean(pull(data_frame, 1)) #mean of numeric vector
mean(pull(data_frame, 2)) #mean of logical vector
mean(pull(data_frame, 3)) #mean of character vector
mean(pull(data_frame, 4)) #mean of factor vector
```
The mean function worked on the numeric and logical vectors. It didn't work of the character and factor vectors as NA was returned. 

### changing to numeric
```{r, eval=FALSE}
as.numeric(pull(data_frame, 2)) #changing the logical vector to numeric
as.numeric(pull(data_frame, 3)) #changing the character vector to numeric
as.numeric(pull(data_frame, 4)) #changing the factor vector to numeric
```
The as.numeric() function worked on the logical vector and the factor vector. For logical vectors the function converts 'TRUE' to '1' and 'FALSE' TO '0', respectively, which explains why the mean() function worked. For the factor vector the as.numeric() function assigns each factor level a number and converts the respective level to the respective number. 

## Problem 2

```{r}
install.packages("palmerpenguins")
```

```{r}
data("penguins", package = "palmerpenguins")
```

### description of dataset
```{r, include=FALSE} 
#include=F because it is just for me to see the df and look at values to include in the description
head(penguins)
penguins$species
penguins$island
```
The penguins dataset contains `nrow(penguins)` rows and `ncol(penguins)` columns. Some of the important variables include the species, island, bill length and depth, flipper length, the body mass and the sex. The islands included in this dataset are "Torgensen," "Dream," and "Biscoe." The species included are "Adelie," "Chinstrap," and "Gentoo." The mean flipper length of penguins in this dataset is `mean(penguins$flipper_length_mm, na.rm=T)`

### scatterplot
```{r}
ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species)) + geom_point()
ggsave(file="penguins_scatterplot.png")
```

